---
title: "figure_5"
author: "Guillaume Lobet & and Ruben Rellan"
date: "February 4, 2015"
output: pdf_document
---

```{r}
library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(magrittr)
library(png)
library(grid)
library(plotly)

```



Plotting reporter data from the ZAT12:LUC2 salt addition experiment


```{r}
# Load the reporter data
reporter <- read.csv("data/reporter.csv")

# Load the segment data
segments <- read.csv("data/seg.csv")
# Use for each time point the complete structure (from time point 1)
seg <- segments[segments$time_in_serie == 1,]
segments <- seg
for(i in 2:max(reporter$time_in_serie)){
  seg$time_in_serie <- i
  segments <- rbind(segments, seg)
}

#calculate growth data

reporter$growth = 0
reporter$growth.x = 0
reporter$growth.y = 0
for(n in reps){
temp <- reporter[reporter$rep_id == n,]
reporter$growth.x[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_x)))
reporter$growth.y[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_y)))
}
reporter$growth <- sqrt(reporter$growth.x^2 + reporter$growth.y^2)
reps <- unique(reporter$rep_id)
reporter$growth = 0
reporter$growth.x = 0
reporter$growth.y = 0
for(n in reps){
temp <- reporter[reporter$rep_id == n,]
reporter$growth.x[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_x)))
reporter$growth.y[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_y)))
}
reporter$growth <- sqrt(reporter$growth.x^2 + reporter$growth.y^2)
reporter <- reporter[reporter$growth < 30,]


reporter1 <- reporter[,c(2,4:9)]
reporter1$diameter <- round(reporter1$diameter, 2)
reporter1$intensity <- round(reporter1$intensity, 2)
reporter1$intensity_ratio <- round(reporter1$intensity_ratio, 2)
reporter1$growth <- round(reporter1$growth , 2)
write.csv(reporter1, "~/Dropbox/research/scripts/websites/gloroot_d3/reporters.csv", row.names=F)



# transform pixels into cm:

reporter <- reporter %>%
  mutate(depth = rep_y/138.6, width=rep_x/138,6)

segments <- segments %>%
  mutate(start_depth = start_y/138.6, start_width=start_x/138.6, end_depth = end_y/138.6, end_width=end_x/138.6)


ggplot(reporter, aes(x = rep_x, y = desc(rep_y), size = intensity)) +

# Add the segment data

geom_segment(data=segments,aes(data=segments,aes(x = start_width, y = -start_depth, xend = end_width, yend = -end_depth), size=0.5, colour="grey", show_guide=F) + facet_grid(~ time_in_serie) +

# Add the reporter data

  geom_point(colour = "magenta", alpha=0.7) +
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggsave("zat12_intensity.pdf", height = 12, width = 8, units = "cm", 
         useDingbats = FALSE)

```


Now we will plot the growth data wit some intensity as an aesthetic depending on zat12 intensity

```{r}


p <- ggplot(reporter, aes(x = width, y = desc(depth), colour = intensity)) +

# Add the segment data

geom_segment(data=segments,aes(x = start_width, y = -start_depth, xend = end_width, yend = -end_depth), size=0.5, colour="grey", show_guide=F) +
facet_grid(~ time_in_serie) +

# Add the reporter data

geom_point() +
scale_colour_gradientn(colours=rainbow(6))+
coord_fixed(ratio = 1) +
theme_bw() +
  ggsave("growth_intensity.pdf", height = 20, width = 22, units = "cm", 
         useDingbats = FALSE)


```

r <- py$ggplotly(p)

Correlation between growth and zat12 expression?

```{r}
reporter %>%
  filter(growth != 0)  %>%
ggplot(aes(x = growth, y = intensity, alpha= time_in_serie)) +
  geom_point(size = 3) +
  theme_bw() +
  ggsave("growth_intensity_correlation.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

```


Root-root interactions data

```{r}
root_root <- read.csv("data/root-root.csv")

p1 <-root_root %>%
        filter(genotype == "Col-LUC2" | genotype == "Sha-LUC2") %>%
        ggplot(aes(x = factor(plants), y = rsa_area, color = factor(plants))) +
        geom_boxplot() +
        geom_point(aes(size =3)) +
        facet_grid(~genotype) +
        labs(x = "Number of plants in rhizotron", y = "Root system area (cm2)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() 
ggsave("r_r_root_area.pdf", height = 8, width = 12, units = "cm", 
         useDingbats = FALSE)

p2 <-root_root %>%
        filter(genotype == "Col-LUC2" | genotype == "Sha-LUC2") %>%
        ggplot(aes(x = factor(plants), y = shoot_area, color = factor(plants))) +
        geom_boxplot() +
        geom_point(aes(size =3)) +
        facet_grid(~genotype) +
        labs(x = "Number of plants in rhizotron", y = "Shoot area (cm2)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
ggsave("r_r_shoot_area.pdf", height = 8, width = 12, units = "cm", 
         useDingbats = FALSE)

grid.arrange(p1, p2, ncol=1)

ggsave("r_r_combined_area.pdf", height = 10, width = 12, units = "cm", 
         useDingbats = FALSE)
        
```


SR9 Experiment

Plotting reporter data from the ZAT12:LUC2 salt addition experiment


```{r}
# Load the reporter data
reporter <- read.csv("data/sr10-reporter.csv")

# Load the segment data
segments <- read.csv("data/sr10-seg.csv")
# Use for each time point the complete structure (from time point 1)
seg <- segments[segments$time_in_serie == 1,]
segments <- seg
for(i in 2:max(reporter$time_in_serie)){
  seg$time_in_serie <- i
  segments <- rbind(segments, seg)
}

#calculate growth data

reporter$growth = 0
reporter$growth.x = 0
reporter$growth.y = 0
for(n in reps){
temp <- reporter[reporter$rep_id == n,]
reporter$growth.x[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_x)))
reporter$growth.y[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_y)))
}
reporter$growth <- sqrt(reporter$growth.x^2 + reporter$growth.y^2)
reps <- unique(reporter$rep_id)
reporter$growth = 0
reporter$growth.x = 0
reporter$growth.y = 0
for(n in reps){
temp <- reporter[reporter$rep_id == n,]
reporter$growth.x[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_x)))
reporter$growth.y[reporter$rep_id == n] <- c(0, abs(diff(temp$rep_y)))
}
reporter$growth <- sqrt(reporter$growth.x^2 + reporter$growth.y^2)
reporter <- reporter[reporter$growth < 30,]


reporter1 <- reporter[,c(2,4:9)]
reporter1$diameter <- round(reporter1$diameter, 2)
reporter1$intensity <- round(reporter1$intensity, 2)
reporter1$intensity_ratio <- round(reporter1$intensity_ratio, 2)
reporter1$growth <- round(reporter1$growth , 2)
write.csv(reporter1, "reporters.csv", row.names=F)



# transform pixels into cm:

reporter <- reporter %>%
  mutate(depth = rep_y/138.6, width=rep_x/138,6)

segments <- segments %>%
  mutate(start_depth = start_y/138.6, start_width=start_x/138.6, end_depth = end_y/138.6, end_width=end_x/138.6)


ggplot(reporter, aes(x = rep_x, y = desc(rep_y), size = intensity)) + 
  geom_segment(data =segments, aes(x = start_width, y = -start_depth, xend = end_width, yend = -end_depth), size=0.5, colour="grey", show_guide=F) +                                                                                      facet_grid(~ time_in_serie) +
  geom_point(colour = "magenta", alpha=0.7) +
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggsave("zat12_intensity.pdf", height = 12, width = 8, units = "cm", 
         useDingbats = FALSE)

```


Now we will plot the growth data wit some intensity as an aesthetic depending on zat12 intensity

```{r}


p <- ggplot(reporter, aes(x = width, y = desc(depth), colour = intensity)) +

# Add the segment data

geom_segment(data=segments,aes(x = start_width, y = -start_depth, xend = end_width, yend = -end_depth), size=0.5, colour="grey", show_guide=F) +
facet_grid(~ time_in_serie) +

# Add the reporter data

geom_point() +
scale_colour_gradientn(colours=rainbow(6))+
coord_fixed(ratio = 1) +
theme_bw() +
  ggsave("growth_intensity.pdf", height = 20, width = 22, units = "cm", 
         useDingbats = FALSE)


```

comparison of root intensity profiles along lateral roots


```{r}

root_intensity <- read.csv("data/reporter_intensity.csv")

root_intensity %>%
  ggplot(aes(x = Distance_From_Tip, y = Intensity, colour = Reporter)) +
  geom_point(size = 3, alpha = 0.6) +
  facet_wrap(~ Reporter, scales="free") +
  theme_bw() +
  ggsave("reporter_intensity.pdf", height = 10, width = 18, units = "cm", 
         useDingbats = FALSE)

```

