---
title: "figure_5_6_7"
author: "Rubén Rellán-Ávarez & Guillaume Lobet"
output: html_document
---

```{r}

library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(magrittr)
library(gridExtra)
library(rmarkdown)
library(hexbin)
library(cowplot)


```

directionality 


```{r}
dir <- read.csv("data/root-data-dir.csv")
dir$angle <- abs(dir$angle)

dir_tidy <- dir %>%
        separate(image, into = c("experiment", "das", "accession", "water", 
                                 "temp", "transfer", "plant"), sep = "-")


ggplot(dir_tidy, aes(x = angle, y = count, colour = water)) +
       facet_grid(temp ~ accession) +
        geom_point(size = 3, alpha = 0.4) +
        geom_smooth()+
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality.pdf", height = 12, width = 16, units = 
                       "cm",useDingbats = FALSE) 

dir_tidy %>%
        filter(accession == "miz1") %>%
  ggplot(aes(x = angle, y = count, colour = water )) +
        geom_point(size = 3, alpha = 0.4) +
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality.pdf", height = 12, width = 16, units = 
                       "cm",useDingbats = FALSE) 

# without the ww high temperature

dir_s <- read.csv("data/root-data-dir-simple.csv")
dir_s$angle <- abs(dir_s$angle)

dir_s_tidy <- dir_s %>%
        separate(image, into = c("experiment", "accession", "water", "temp", "das", "plant"), sep = "-")


ggplot(dir_s_tidy, aes(x = angle, y = count, colour = water)) +
       facet_grid(~ accession) +
        geom_point(size = 3, alpha = 0.4) +
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality.pdf", height = 12, width = 16, units = 
                       "cm",useDingbats = FALSE) 

dir_hormone <- read.csv("data/root-data-sobel-dir.csv")
dir_hormone$angle <- abs(dir_hormone$angle)

dir_hormone_tidy <- dir_hormone %>%
        separate(image, into = c("experiment", "das", "accession", "water",
                                "plant", "aq"), sep = "-")

ggplot(dir_hormone_tidy, aes(x = angle, y = count, colour = water)) +
       facet_grid(~ accession) +
        geom_point(size = 3, alpha = 0.5) +
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality.pdf", height = 12, width = 16, units = 
                       "cm",useDingbats = FALSE) 


dir_half <- read.csv("data/root-data-half-dir.csv")
dir_half$angle <- abs(dir_half$angle)

dir_half_tidy <- dir_half %>%
        separate(image, into = c("experiment", "das", "accession", "water",
                                "plant", "aq"), sep = "-")

ggplot(dir_half_tidy, aes(x = angle, y = count, colour = water)) +
       facet_grid(~ accession) +
        geom_point(size = 3, alpha = 0.5) +
        geom_smooth() +
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality.pdf", height = 12, width = 16, units = 
                       "cm",useDingbats = FALSE) 


dir_hormone_ks <- read.csv("data/root-data-half-dir-ks.csv")
dir_hormone_ks$angle <- abs(dir_hormone_ks$angle)

              
ks.test(dir_hormone_ks$miz_ww, dir_hormone_ks$miz_wd)


# experiment 64

dir64 <- read.csv("data/64-root-data-dir.csv")
dir64$angle <- abs(dir64$angle)

dir_tidy64 <- dir64 %>%
        separate(image, into = c("experiment", "temp", "water", "transfer", 
                                 "das", "plant", "extra"), sep = "-")
dir_tidy64 <- tbl_df(dir_tidy64)

dir_tidy64 %>%
  filter(transfer == "d9") %>%
  ggplot(aes(x = angle, y = count, colour = water)) +
       facet_grid(temp ~ transfer) +
        geom_smooth()+
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("directionality64.pdf", height = 12, width = 12, units = 
                       "cm",useDingbats = FALSE) 




# time course experiment


dir_tc <- read.csv("data/time-course-root-data-dir.csv")
dir_tc$angle <- abs(dir_tc$angle)

dir_tc_tidy <- dir_tc %>%
        separate(image, into = c("experiment", "das", "accesion", "water", 
                                 "plant", "extra"), sep = "-")
dir_tc_tidy <- tbl_df(dir_tc_tidy)

ggplot(dir_tc_tidy, aes(x = angle, y = count, colour = water)) +
       facet_grid(~ das ) +
        geom_smooth()+
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()



# time course experiment with depth data

dir_tc_depth <- read.csv("data/time-course-depth-root-data-dir.csv")
dir_tc_depth$angle <- abs(dir_tc_depth$angle)

dir_tc_depth_tidy <- dir_tc_depth %>%
        separate(image, into = c("experiment", "das", "accesion", "water", 
                                 "plant", "extra"), sep = "-")
dir_tc_depth_tidy <- tbl_df(dir_tc_depth_tidy)

dir_tc_depth_tidy %>%
  filter(depth < 10) %>%
  ggplot(aes(x = angle, y = count, colour = water)) +
      geom_point(size = 2, alpha = 0.2) +
      ylim(0, 4) +
      facet_grid(depth ~ das) +
      geom_smooth()+
      labs(x = "Angle", y = "count") +
      theme(panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())

````

Script for the plot of the relationship between the soil water content and the root angle



## Complete images

First we try to plot the relation between the soil water content using **complete images** of the root segment. transformation between pixel values and humidity is based on the regression plot found the figure 5 repo.

```{r}
seg <- read.csv("data/2015-02-23-soil-data-all.csv")
seg$humidity <- 221.0 - (3.96 * seg$environment)
seg$das <- substr(seg$image, start = 4, stop = 5)
seg$treatment <- substr(seg$image, start = 11, stop = 12)

# Create a depth factor
max <- max(seg$start_y)
seg$depth[seg$start_y < (max/3)] <- 1
seg$depth[seg$start_y > (max/3) & seg$start_y < (2*(max/3))] <- 2
seg$depth[seg$start_y > (2*(max/3))] <- 3

```


First we create a faceted plot, byt das and depth

```{r}

# Create a faceted plot by time and depth
ggplot(data = seg, aes(x = humidity, y = direction)) + 
  facet_grid(depth~das) +
  geom_point(aes(colour=treatment), size=1, alpha=0.7) + 
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") 

```

then we look at the regression on the whole dataset. It does not look that good for me....

```{r}

# Plot the regression for the whole dataset
fit <- lm(seg$direction ~ seg$humidity)
summary(fit)

ggplot(data = seg, aes(x = humidity, y = direction)) + 
  geom_abline(intercept = fit$coefficients[[1]], slope = fit$coefficients[[2]], colour="grey")+
  geom_point(aes(colour=treatment), size=1, alpha=0.7) + 
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") 

```

Then we try a regression on mean values (by das). It look a bit better.

```{r}
# Plot the regression for the mean values
rs <- ddply(seg, .( treatment, das), summarise, 
            dir = mean(direction), 
            dir.sd = sd(direction), 
            envi=mean(humidity, na.rm = T),
            envi.sd=sd(humidity, na.rm = T))

fit <- lm(rs$dir ~ rs$envi)
print(summary(fit))

ggplot(data = rs,aes(x = envi, y = dir)) + 
  geom_abline(intercept = fit$coefficients[[1]], slope = fit$coefficients[[2]], colour="grey")+
  geom_point(aes(colour=treatment), size=2, alpha=0.7) + 
  geom_errorbar(aes(ymin = dir-dir.sd, ymax = dir+dir.sd, height=0, colour=treatment), alpha=0.7) + 
  geom_errorbarh(aes(xmin = envi-envi.sd, xmax = envi+envi.sd, height=0, colour=treatment), alpha=0.7) +
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") 


````

## Differential images

Next we try to plot the relation between the soil water content but using **differential images** images of the root segment

```{r}

seg <- read.csv("data/2015-02-23-soil-data-diff.csv")
seg$humidity <- 221.0 - (3.96 * seg$environment)
seg$das <- substr(seg$image, start = 4, stop = 5)
seg$treatment <- substr(seg$image, start = 11, stop = 12)

# Create a depth factor
max <- max(seg$start_y)
seg$depth[seg$start_y < (max/3)] <- 1
seg$depth[seg$start_y > (max/3) & seg$start_y < (2*(max/3))] <- 2
seg$depth[seg$start_y > (2*(max/3))] <- 3

```


First we create a faceted plot, by das and depth

```{r}
# Create a faceted plot by time and depth

seg %>%
  filter(das >15) %>%
  ggplot(aes(x = humidity, y = direction)) + 
  facet_grid(~das) +
  geom_point(aes(colour=treatment), size=3, alpha=0.7) + 
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") 


```


Next, we make a regression on the whole dataset. Still does not look good

```{r}
# Plot the regression for the whole dataset
fit <- lm(seg$direction ~ seg$humidity)
print(summary(fit))

ggplot(data = seg, aes(x = humidity, y = direction)) + 
  geom_abline(intercept = fit$coefficients[[1]], slope = fit$coefficients[[2]], colour="grey")+
  geom_point(aes(colour=treatment), size=1, alpha=0.7) +
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") 

# Plot the regression by day
fit <- lm(seg$direction ~ seg$humidity)
print(summary(fit))

ggplot(data = seg, aes(x = humidity, y = direction)) +
  geom_point(aes(colour=treatment), size=1, alpha=0.7) +
  facet_grid(depth~das) +
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") 


```


Then we try a regression in the mean data (by das). This is the best one for me...

```{r}
# Plot the regression for the mean values
rs <- ddply(seg, .( treatment, das), summarise, 
            dir = mean(direction), 
            dir.sd = sd(direction), 
            envi=mean(humidity, na.rm = T),
            envi.sd=sd(humidity, na.rm = T)),
            dir.se = se(direction), 
            envi.se=se(humidity, na.rm = T))

fit <- lm(rs$dir ~ rs$envi)
print(summary(fit))

ggplot(data = rs,aes(x = envi, y = dir)) + 
  geom_point(aes(colour=treatment), size=2, alpha=0.7) + 
  geom_errorbar(aes(ymin = dir-dir.sd, ymax = dir+dir.sd, height=0, colour=treatment), size=0.5, alpha=0.7) + 
  geom_errorbarh(aes(xmin = envi-envi.sd, xmax = envi+envi.sd, height=0, colour=treatment), size=0.5, alpha=0.7) +
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") +
    ggsave("angle_moisture.pdf", height = 8, width = 10, units = 
                       "cm",useDingbats = FALSE) 
```
# Alternative graphs 

```{r}
# simple graph
ggplot(data = seg, aes(x = humidity, y = direction)) +
  geom_point(aes(colour=treatment), size=3, alpha=0.7) +
  xlim(30, 80)+
  theme_bw()+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") +
  ggsave("angle_moisture_no_average.pdf", height = 8, width = 10, units = 
                       "cm",useDingbats = FALSE)

# rug graph

ggplot(data = seg, aes(x = humidity, y = direction)) +
  geom_point(aes(colour=treatment), size=3, alpha=0.7) +
  xlim(30, 80)+
  theme_bw()+
  geom_rug(col=rgb(.5,0,0,alpha=.2))+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") +
  ggsave("angle_moisture_rug.pdf", height = 8, width = 10, units = 
                       "cm",useDingbats = FALSE)

# rug graph with facets
ggplot(data = seg, aes(x = humidity, y = direction)) +
  geom_point(aes(colour=treatment), size=3, alpha=0.7) +
  xlim(30, 80)+
  theme_bw()+
  geom_rug(col=rgb(.5,0,0,alpha=.2))+
  facet_grid(~treatment)+hi
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") +
  ggsave("angle_moisture_rug_facet.pdf", height = 8, width = 10, units = 
                       "cm",useDingbats = FALSE)

# with geom density

ggplot(data = seg, aes(x = humidity, y = direction)) +
  geom_point(aes(colour=treatment), size=3, alpha=0.7) +
  xlim(30, 80)+
  theme_bw()+
  geom_density2d(aes(color=treatment))+
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") +
  ggsave("angle_moisture_density2d.pdf", height = 8, width = 10, units = 
                       "cm",useDingbats = FALSE)

# with histograms on top and sides

scatter <- ggplot(data = seg, aes(x = humidity, y = direction)) +
  geom_point(aes(colour=treatment), size=3, alpha=0.7) +
  xlim(30, 85)+
  theme(legend.position=c(1,1),legend.justification=c(1,1)) +
  ylab("Root direction [°]") +
  xlab("Moisture content [%]") 

histogram_x <- ggplot(data = seg, aes(humidity)) +
  geom_density(aes(colour = treatment, fill = treatment, alpha = 0.7)) +
  xlim(30, 80) +
  theme(legend.position = "none") 

histogram_y <- ggplot(data = seg, aes(direction)) +
  geom_density(aes(colour = treatment, fill = treatment, alpha = 0.7)) +
  coord_flip() +
  theme(legend.position = "none") 

empty <- ggplot()+geom_point(aes(1,1), colour="white") +
     theme(                              
       plot.background = element_blank(), 
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.border = element_blank(), 
       panel.background = element_blank(),
       axis.title.x = element_blank(),
       axis.title.y = element_blank(),
       axis.text.x = element_blank(),
       axis.text.y = element_blank(),
       axis.ticks = element_blank()
     ) +

grid.arrange(histogram_x, empty, scatter, histogram_y, ncol=2, nrow=2, widths=c(4, 1), heights=c(1, 4)) +
  ggsave("angle_moisture_histogram_scatter.pdf", height = 10, width = 10, units = 
                       "cm",useDingbats = FALSE)
  

# with histograms on top and sides and geom_density

scatter <- ggplot(data = seg, aes(x = direction, y = humidity)) +
  geom_point(aes(colour=treatment), size=2, alpha=0.7) +
  ylim(35, 80)+
  geom_density2d(aes(color=treatment))+
  ylab("Moisture content [%]") + 
  xlab("Root direction [°]") +
  theme_bw() +
  theme(legend.position=c(1,1),legend.justification=c(1,1))

histogram_x <- ggplot(data = seg, aes(direction)) +
  geom_density(aes(colour = treatment, fill = treatment, alpha = 0.7)) +
  xlab("Root Direction [º]") +
  theme_bw() +
  theme(legend.position = "none") 

histogram_y <- ggplot(data = seg, aes(humidity)) +
  geom_density(aes(colour = treatment, fill = treatment, alpha = 0.7)) +
  xlab("Moisture content [%]") +
  xlim(35, 80) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none") 

empty <- ggplot()+geom_point(aes(1,1), colour="white") +
     theme(                              
       plot.background = element_blank(), 
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.border = element_blank(), 
       panel.background = element_blank(),
       axis.title.x = element_blank(),
       axis.title.y = element_blank(),
       axis.text.x = element_blank(),
       axis.text.y = element_blank(),
       axis.ticks = element_blank()
     )
blankPanel<-grid.rect(gp=gpar(col="white"))

grid.arrange(histogram_x, blankPanel, scatter, histogram_y, ncol=2, nrow=2, widths=c(3.5, 1), heights=c(1, 3.5)) +
  ggsave("angle_moisture_histogram_scatter_density.pdf", height = 10, width = 10, units = "cm",useDingbats = FALSE)

```

# relative water content

```{r}

rwc <- read.csv("data/rwc.csv")
rwc %>%
  ggplot(aes(x = transfer, y = RWC, color = water)) +
  geom_boxplot() +
  xlab("Transfer Day") +
  ylab("Relative water content (%)") +
  facet_grid(~temp) +
  theme_bw()
  ggsave("rwc.pdf", height = 10, width = 10, units = "cm", useDingbats = FALSE)

```

# moisture calibration

```{r}

moist_cal <- read.csv("data/moisture_calib.csv")
moist_cal %>%
  ggplot(aes(x = pixel_int, y = moisture)) +
  geom_point(size = 3) +
  stat_smooth(method=lm) +
  xlab("Pixel intensity") +
  ylab("Water content (%)") +
  draw_text("r2 = 0.85", angle = 0, size = 12, alpha = .8)
  ggsave("moisture_calibration.png", height = 10, width = 10, units = "cm")

```

# root brightness

```{r}

root_brightness <- read.csv("data/root-brightness.csv")
root_brightness %>%
  ggplot(aes(x = water, y = intensity, colour = root_presence)) +
  geom_boxplot() +
  geom_point(size = 3, alpha = 0.7, position=position_jitter(w = 0.5)) +
  labs(colour = "Root Presence") +
  xlab("Water treatment") +
  ylab("Intensity") +
  ggsave("root_brightness.png", height = 10, width = 10, units = "cm")
  
root_brightness %>%
wilcox.test(intensity ~ root_presence , data = root_brightness)

root_WD <- root_brightness %>%
  filter (water == "WD")
wilcox.test(intensity ~ root_presence , data = root_WD)

root_WW <- root_brightness %>%
  filter (water == "WW")
wilcox.test(intensity ~ root_presence , data = root_WW)

```

Redrawing the directionality data:

```{r}
dir_data <- read.csv("data/figure_6_A-D.csv")

dir_short <- dir_data %>%
        filter(transfer == "13d") %>%
        filter(accession != "sha")

ggplot(dir_short, aes(x = angle, y = count, colour = water)) +
        facet_grid(.~ accession) +
        ylim(0, 2.5) +
        geom_smooth() +
        labs(x = "Angle", y = "count") +
        theme_bw() +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality.pdf", height = 12, width = 16, units = 
                       "cm",useDingbats = FALSE) 


```

