---
title: "figure_8"
author: "Rubén Rellán-Ávarez"
output: html_document
---

library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(magrittr)
library(gridExtra)
library(rmarkdown)


It does seem that the width of the low phosphorus plants is larger they are 
shorter. Let´s take a look at the global data.

```{r}
global <- read.csv("data/root-data-global.csv")

# some data tyding 

global_tidy <- global %>%
        separate(image, into = c("experiment", "das", "phosphorus", "plant", 
                                 "aq"), sep = "-")

ggplot(global_tidy, aes(x = phosphorus, y = width, colour = phosphorus)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "Phosphorus availability", y = "Width (cm)") +
        facet_grid(das ~.) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

ggsave("width.pdf", height = 10, width = 10, units = "cm", 
         useDingbats = FALSE)

aov.width <- aov(width ~ phosphorus * das, data = global_tidy)
summary(aov.width)

```


However it seems that the area is pretty similar. Let´s check that and also the 
depth, and AR.

```{r}

# area

ggplot(global_tidy, aes(x = phosphorus, y = convexhull, colour = phosphorus)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "Phosphorus availability", y = "Area (cm2)") +
        facet_grid(das ~.) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("area.pdf", height = 14, width = 10, units = "cm", 
         useDingbats = FALSE)

aov.area <- aov(convexhull ~ phosphorus * das, data = global_tidy)
summary(aov.area)

# depth

ggplot(global_tidy, aes(x = phosphorus, y = depth, colour = phosphorus)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "Phosphorus availability", y = "Depth (cm)") +
        facet_grid(das ~.) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("depth.pdf", height = 14, width = 12, units = "cm", 
         useDingbats = FALSE)

aov.depth <- aov(depth ~ phosphorus * das, data = global_tidy)
summary(aov.depth)

# Aspect ratio

ggplot(global_tidy, aes(x = phosphorus, y = ar, colour = phosphorus)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "Phosphorus availability", y = "Depth/width") +
        facet_grid(das ~.) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("ar.pdf", height = 14, width = 10, units = "cm", 
         useDingbats = FALSE)

aov.ar <- aov(ar ~ phosphorus * das, data = global_tidy)
summary(aov.ar)

```

I will now take a look at root-shoots correlations.

```{r}

# shoot/root correlations (areas)

ggplot(global_tidy, aes(x = convexhull, y = shoot_area, colour = phosphorus)) +
        geom_point(size = 4) +
        labs(x = "Root Area (cm2)", y = "Shoot Area (cm2)") +
        facet_grid(das ~.) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("root-shoot-correlation.pdf", height = 14, width = 10, units = 
                       "cm", useDingbats = FALSE)

```


I don´t see very apparent changes in term of angles. Let´s see. 


```{r}
dir <- read.csv("data/root-data-dir.csv")
dir$angle <- abs(dir$angle)

dir_tidy <- dir %>%
        separate(image, into = c("experiment", "das", "phosphorus", "plant", 
                                 "aq"), sep = "-")

ggplot(dir_tidy, aes(x = angle, y = count, colour = phosphorus)) +
        labs(x = "Angle", y = "count") +
        geom_smooth() +
        facet_grid(das ~.) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("directionality-P.pdf", height = 15, width = 11, units = 
                       "cm", useDingbats = FALSE) 


# Kolmogorov-Smirnoff analysis of the directionality distributions

dir_P_ks <- read.csv("data/root-data-dir-P-ks.csv")

ks.test(dir_P_ks$early_high, dir_P_ks$early_low)
ks.test(dir_P_ks$late_high, dir_P_ks$late_low)

```
