---
title: "figure3"
author: "Rubén Rellán Álvarez"
date: "December 26, 2014"
output: html_document
---

```{r}
library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(magrittr)
library(gridExtra)
library(stringr)

```

Global analysis of example of root growth. We include in this file the manual measurements.
`
```{r}

global <- read.csv("data/root-data-global.csv")
global <- tbl_df(global)


```

We will first do some correlation analysis between the automatic measurements and the manual quantification

```{r}

# Depth can be a proxy for PR length
ggplot(global, aes(x=PrPathLength, y=depth))  +
        geom_point(size=3) +
        geom_smooth(method = "lm", colour = "grey60") +
        labs(x = "PR Lenght Manual (cm)", y = "Depth automatic (cm)")
ggsave("9_PRPath_DepthCorr.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

modeldepth <- lm(depth ~ PrPathLength, global)
summary(modeldepth)


# Convex Hull as a proxy for Root Systems area
ggplot(global, aes(x=AreaManual, y=convexhull))  +
        geom_point(size=3) +
        geom_smooth(method = "lm", colour = "grey60") +
        labs(x = "Area Manual (cm2)", y = "Area automatic (cm2)")
ggsave("9_AreaManual_AutomCorr.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

modelarea <- lm(convexhull ~ AreaManual, global)
summary(modelarea)

```

Nice, now, let´s plot some data


```{r}
# PR Growth Rate

p1 <- ggplot(global, aes(x=das, y=PrGrowthRate)) + 
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "PR Growth rate (cm/day)") +
        theme_bw() +
ggsave("9_PRGrowthRate.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)


# Depth

p2 <-ggplot(global, aes(x=das, y=desc(depth))) +
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "Depth(cm)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("9_depth.pdf", width = 8, height = 8, units = "cm", useDingbats = FALSE)

# Root System Area

p3 <- ggplot(global, aes(x=das, y=AreaManual)) + 
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "Area(cm2)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("9_area.pdf", width = 8, height = 8, units = "cm", useDingbats = FALSE)


# Width
p4 <- ggplot(global, aes(x=das, y=width)) +
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "Width(cm)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("9_width.pdf", width = 8, height = 8, units = "cm", useDingbats = FALSE)

# Aspect ratio
p5 <- ggplot(global, aes(x=das, y=ar))  + 
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "depth/width") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("9_ar.pdf", width = 8, height = 8, units = "cm", useDingbats = FALSE)

# lateral root number
p6 <- ggplot(global, aes(x=das, y=LrNumber) ) +
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "Lateral root number") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("9_LR#.pdf", width = 8, height = 8, units = "cm", useDingbats = FALSE)

# branched zone area
p7 <- ggplot(global, aes(x=das, y=BranchedZone)) + 
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "Branched Zone Area") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("9_branched_zone.pdf", width = 8, height = 8, units = "cm", useDingbats = FALSE)

# LR density in PR
p8 <- ggplot(global, aes(x=das, y=LRDensityPrPath) )  +
        geom_point(size=3, colour = "grey20") +
        labs(x = "Days after Sowing", y = "# LR / PR length")+ 
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw() +
        ggsave("9_LR_density.pdf", width = 8, height = 8, units = "cm", useDingbats = FALSE)

grid.arrange(p1, p2, p4, p3, p8, nrow = 1) +
        ggsave("combined.pdf", width = 14, height = 8, units = "cm", useDingbats = FALSE)

```

I will now do some analysis on the directionality data.

```{r}

dir <- read.csv("data/root-data-dir.csv")
dir$angle <- abs(dir$angle)

dir_tidy <- dir %>%
        separate(image, into = c("experiment", "plant", "das", sep="_"))

ggplot(dir_tidy, aes(x=angle, y=count)) +
        geom_smooth() +
  facet_grid (~das) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
theme_bw()
        ggsave("directionality_line.pdf", width = 14, height = 5 , units = "cm", useDingbats = FALSE)

```

              
new GLO-RIA version with depth data  

````{r}
dir3 <- read.csv("data/root-data-dir2.csv")
dir3$angle <- abs(dir3$angle)

dir3_tidy <- dir3 %>%
        separate(image, into = c("experiment", "plant", "das", sep="_"))

ggplot(dir3_tidy, aes(x=angle, y=count, fill = y))+
  geom_point(size=2, aes(alpha = 0.8)) +
        ylim(0, 6) +
  facet_grid (y ~ das) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())
 
```

new GLO-RIA version with fourier

````{r}

dir_fou <- read.csv("data/root-data-fou-dir.csv")
dir_fou$angle <- abs(dir_fou$angle)

dir_fou_tidy <- dir_fou %>%
        separate(image, into = c("experiment", "plant", "das", sep="_"))

ggplot(dir_fou_tidy, aes(x=angle, y=count)) +
  geom_point(size=3, aes(alpha = 0.5)) +
        ylim(0, 6) +
  facet_grid (~das) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

```

Testing out GLO_RIA´s new version of local.

```{r}

dir_local <- read.csv("data/root-data-local.csv")
dir_local$angle <- abs(dir_local$angle)

dir_local_tidy <- dir_local %>%
        separate(image, into = c("experiment", "plant", "das", sep="_"))

dir_local_tidy[,4] <- NULL

ggplot(dir_local_tidy, aes(x=angle, fill = plant))+
  geom_density(alpha=.5) +
  facet_grid (~ das) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

ggplot(dir_local_tidy, aes(x=x, y= desc(y), colour = angle ))+
  geom_point(alpha=0.5) +
  facet_grid (~ das) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

```


