---
title: "Validation of GLO-RIA"
author: Rubén Rellán-Álvarez, Guillaume Lobet, Heike Lindner, Pierre-Luc Pradier,
  Jose Sebastian1, Muh-Ching Yee, Yu Geng, Charlotte Trontin, Therese LaRue, Amanda
  Schrager, Cara Haney, Rita Nieu, Julin Maloof, John P. Vogel, José R. Dinneny
date: "17 Jul 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

This document presents the validation of different GLO-RIA measurments. Two aproaches were chosen for the validation. First, we compared some of the measurements from GLO-RIA with measurements done manually. Second, we used the ArchiSimple model (Pagès et al. 2013) to generated images of root systems. Since the images are created using a model, we know the ground-truth values for each of them. This allows us to validate GLORIA on a large number of ground-truth images.

```{r lib_loading, message=F, echo=F, warning=F}
library(plyr)
library(ggplot2)
library(gridExtra)
library(data.table)
library(Hmisc)
```


# Validation using manual measurements of real images

This validation was carried out in an independent set of 15 images of different arabidopsis accesions grown in control conditions.

Root system depth and width and individual lateral root angles were measured in the images and then GLO-RIA was used to measure the same parameters.

## Directionality

```{r, fig.width=15}

man_gloria <- read.csv("manual_gloria.csv")

fit <- lm(man_gloria$man_dir ~ man_gloria$gloria_dir)
r2 <- round(summary(fit)$r.squared, 3)
sp <- round(rcorr(man_gloria$gloria_dir,man_gloria$man_dir, type="spearman")$r[1,2], 3)
pe <- round(rcorr(man_gloria$gloria_dir,man_gloria$man_dir, type="pearson")$r[1,2], 3)
leg <- paste("r-squared = ",r2, "\nPearson = ",pe, "\nSpearman = ",sp)

plot_man_dir <- ggplot(man_gloria, aes(gloria_dir, man_dir)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method='lm') +
  xlab("Mean manual directionality [°]") + ylab("Mean estimated directionality [°]") + 
  theme_bw() +
  annotate("text", y= max(man_gloria$man_dir), x =max(man_gloria$gloria_dir), 
           label=leg, hjust=1, vjust=5, size=5) +
  theme(plot.title=element_text(hjust=0, size=30)) + 
  ggsave(file="figures/man-dir.pdf", width=6, height=6)

```

## Depth

```{r, fig.width=15}

fit <- lm(man_gloria$man_d ~ man_gloria$gloria_d)
r2 <- round(summary(fit)$r.squared, 3)
sp <- round(rcorr(man_gloria$gloria_d,man_gloria$man_d, type="spearman")$r[1,2], 3)
pe <- round(rcorr(man_gloria$gloria_d,man_gloria$man_d, type="pearson")$r[1,2], 3)
leg <- paste("r-squared = ",r2, "\nPearson = ",pe, "\nSpearman = ",sp)

plot_man_d <- ggplot(man_gloria, aes(gloria_d, man_d)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method='lm') +
  xlab("Manual depth [cm]") + ylab("Estimated depth [cm]") + 
  theme_bw() +
  annotate("text", y= max(man_gloria$man_d), x =max(man_gloria$gloria_d), 
           label=leg, hjust=1, vjust=5, size=5) +
  theme(plot.title=element_text(hjust=0, size=30)) + 
  ggsave(file="figures/man-depth.pdf", width=6, height=6)

```

## Width

```{r, fig.width=15}

fit <- lm(man_gloria$man_w ~ man_gloria$gloria_w)
r2 <- round(summary(fit)$r.squared, 3)
sp <- round(rcorr(man_gloria$gloria_w,man_gloria$man_w, type="spearman")$r[1,2], 3)
pe <- round(rcorr(man_gloria$gloria_w,man_gloria$man_w, type="pearson")$r[1,2], 3)
leg <- paste("r-squared = ",r2, "\nPearson = ",pe, "\nSpearman = ",sp)

plot_man_w <- ggplot(man_gloria, aes(gloria_w, man_w)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method='lm') +
  xlab("Manual width [cm]") + ylab("Estimated width [cm]") + 
  theme_bw() +
  annotate("text", y= max(man_gloria$man_w), x =max(man_gloria$gloria_w), 
           label=leg, hjust=1, vjust=5, size=5) +
  theme(plot.title=element_text(hjust=0, size=30)) + 
  ggsave(file="figures/man-width.pdf", width=6, height=6)

```

grid.arrange(plot_man_dir, plot_man_d, plot_man_w, ncol=3) 

# Validation using ArchiSimple model generated images

# Directionality

```{r load_dir_data, message=F, echo=F, eval=F}
gloria <- fread("data/gloria-directionality.csv", header = T) # The estimators
rsml <- fread("data/rsml-directionality.csv", header = T) # The estimators
```

For the validation of directionality measurements, we used Archisimple to generate root systems with contrasted gravitropism. We compared the results from GLO-RIA with the ground-truth values from the model.

```{r format_dir_data, message=F, echo=F, eval=F}

gloria$angle <- abs(gloria$angle)

rsml$theta <- rsml$theta * (180 / pi) # Need to format the angle formt he RSML so they can be compared to the angles of GLORIA
rsml$theta <- rsml$theta - 270
rsml <- rsml[rsml$theta <= 90 & rsml$theta >= -90,]
rsml$theta <- abs(rsml$theta)

rsml$image <- as.character(rsml$image)
gloria$image <- as.character(gloria$image)
gloria$image <- substr(gloria$image, 0, nchar(gloria$image)-4)
ims <- unique(rsml$image)
ims <- ims[!is.na(ims)]



# Loop over the parameters space
counter <- 0
tot_sim <- length(ims)
percent <- 0
mean.rs <- data.frame(id=character(),sr=numeric(), gl1=numeric())

# create a new table with the mean value per image
for(i in ims){
  counter <- counter+1
  prog <- ( counter / tot_sim ) * 100
  if(prog > percent){
    message(paste(round(prog), "% of data analysed"))
    percent <- percent + 5
  }
  temp <- gloria[gloria$image == i]
  
  # Get the mean angle from the distribution
  temp <- data.frame(id=i, 
                     sr=mean(rsml$theta[rsml$image == i]),#sum(abs(h2$angle) * h1$count) / sum(h1$count), 
                     gl1=sum(temp$angle * temp$count) / sum(temp$count))  
  mean.rs <- rbind(mean.rs, temp)
}


mean.rs$id <- as.character(mean.rs$id)
for(i in 1:nrow(mean.rs)){
  mean.rs$gravi[i] <- as.numeric(strsplit(mean.rs$id[i], "-")[[1]][6])
  mean.rs$insert[i] <- as.numeric(strsplit(mean.rs$id[i], "-")[[1]][11])
}
# Different insertion angles vwhere used to generate the dataset. We keep only one insertion angle, to have a more uniform dataset.

mean.rs <- mean.rs[mean.rs$insert == 1.5,]

```

```{r load_mean_dir, echo=F}
mean.rs <- read.csv("data/mean.directionality.csv")

```

```{r plot_dir, message=F, echo=F,fig.width=15}

# Remove outliyers
b1 <- boxplot(mean.rs$sr, plot=F)
b2 <- boxplot(mean.rs$gl1, plot=F)
temp <- mean.rs[mean.rs$sr < b1$stats[5,],]


# Look at the relation between the directionality and the root orientation
fit <- lm(temp$gl1 ~ temp$sr)
r2 <- round(summary(fit)$r.squared, 3)
sp <- round(rcorr(temp$sr,temp$gl1, type="spearman")$r[1,2], 3)
pe <- round(rcorr(temp$sr,temp$gl1, type="pearson")$r[1,2], 3)
leg <- paste("r-squared = ",r2, "\nPearson = ",pe, "\nSpearman = ",sp)

plot1 <- ggplot(temp, aes(sr, gl1)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method='lm') +
  ylab("Mean estimated directionality [°]") + 
  xlab("Mean root orientation [°]") + 
  theme_bw() +
  annotate("text", y= min(temp$gl1), x =max(temp$sr), 
           label=leg, hjust=1, vjust=0, size=5) + 
  # ggtitle('A') + 
  theme(plot.title=element_text(hjust=0, size=30))+ 
  ggsave(file="figures/model-dir.pdf", width=6, height=6)



# Look at the relation with the gravitropism (keep only one insertion angle)
fit <- lm(temp$gl1 ~ temp$gravi)
r2 <- round(summary(fit)$r.squared, 3)
sp <- round(rcorr(temp$gravi,temp$gl1, type="spearman")$r[1,2], 3)
pe <- round(rcorr(temp$gravi,temp$gl1, type="pearson")$r[1,2], 3)
leg <- paste("r-squared = ",r2, "\nPearson = ",pe, "\nSpearman = ",sp)

plot2 <- ggplot(temp, aes(gravi, gl1)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method='lm') +
  xlab("Gravitropism strength [-]") + ylab("Mean estimated directionality [°]") + 
  theme_bw() +
  annotate("text", y= max(temp$gl1), x =max(temp$gravi), 
           label=leg, hjust=1, vjust=1, size=5) + 
  #ggtitle('B') + 
  theme(plot.title=element_text(hjust=0, size=30)) + 
  ggsave(file="figures/model-gravi.pdf", width=6, height=6)

plot <- grid.arrange(plot1, plot2, ncol=2)

```

## Size related measurements


```{r load_data_size, echo=FALSE}
gtruth <- fread("data/rsml-size.csv", header = T) 
global <- fread("data/gloria-size.csv", header = T) 
```

For the validation of size variables, we used Archisimple to generate `r nrow(global)` root systems with contrasted size. We compared the results from GLO-RIA with the ground-truth values from the model.

## Depth

```{r plot_size, echo=FALSE, fig.width=15}
temp <- data.frame(x=gtruth$depth, y=global$depth)
fit <- lm(temp$y ~ temp$x)
r2 <- round(summary(fit)$r.squared, 3)
sp <- round(rcorr(temp$x,temp$y, type="spearman")$r[1,2], 3)
pe <- round(rcorr(temp$x,temp$y, type="pearson")$r[1,2], 3)
leg <- paste("r-squared = ",r2, "\nPearson = ",pe, "\nSpearman = ",sp)

plot1 <- ggplot(temp, aes(x, y)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method='lm') +
  geom_abline(intercept = 0, slope = 1, lty=2) + 
  ylab("Estimated root system depth [px]") + 
  xlab("Real root system depth [px]") + 
  theme_bw() +
  annotate("text", y= min(temp$y), x =max(temp$x), 
           label=leg, hjust=1, vjust=0, size=5) + 
  # ggtitle('A') + 
  theme(plot.title=element_text(hjust=0, size=30)) + 
  ggsave(file="figures/model-depth.pdf", width=6, height=6)

```

## Width

```{r plot_width, echo=FALSE, fig.width=15}
temp <- data.frame(x=gtruth$width, y=global$width)
fit <- lm(temp$y ~ temp$x)
r2 <- round(summary(fit)$r.squared, 3)
sp <- round(rcorr(temp$x,temp$y, type="spearman")$r[1,2], 3)
pe <- round(rcorr(temp$x,temp$y, type="pearson")$r[1,2], 3)
leg <- paste("r-squared = ",r2, "\nPearson = ",pe, "\nSpearman = ",sp)

plot2 <- ggplot(temp, aes(x, y)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method='lm') +
  geom_abline(intercept = 0, slope = 1, lty=2) + 
  ylab("Estimated root system width [px]") + 
  xlab("Real root system width [px]") + 
  theme_bw() +
  annotate("text", y= min(temp$y), x =max(temp$x), 
           label=leg, hjust=1, vjust=0, size=5) + 
  # ggtitle('B') + 
  theme(plot.title=element_text(hjust=0, size=30)) + 
  ggsave(file="figures/model-width.pdf", width=6, height=6)
```



# References 

Pagès, L., Bécel, C., Boukcim, H., Moreau, D., Nguyen, C., & Voisin, A.-S. (2013). Calibration and evaluation of ArchiSimple, a simple model of root system architecture, 290, 76–84. http://doi.org/10.1016/j.ecolmodel.2013.11.014