---
title: "figure_3_heritability"
author: "Rubén Rellán-Ávarez", "Guillaume Lobet"
output: html_document
---

```{r}

library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(magrittr)
library(gridExtra)
library(rmarkdown)
library(shapes)
library(geomorph)


```

First let´s take a look at the basic global descriptors

```{r}
global <- read.csv("data/high-root-data-global.csv")

# some data tyding 

global_tidy <- global %>%
        separate(image, into = c("experiment", "das", "accession", "plant", 
                                 "aq"), sep = "-")
```

Now, lets work on the heritability:

```{r}

# Example for width:

aov.width <- aov(width ~ accession, data = global_tidy)
summary(aov.width)

# According to Olivier Loudet, the genetic variance is estimated by:

VG = (Var'Genotype' - Var'Residuals') / n

# where n is the (mean) number of replicate measurements for each genotype in the experiment. Not sure what this could be in our case.

# The 'Residuals' are a good estimator of the environmental variance in the experiment: 

VE = Var"Residuals" / n

# Total phenotipic variance would then be:

VP (Total Phenotypic variance) = VG + VE

# and therefore the (broad sense) heritability would be:

h2 = VG / VP

# So in our case for width, we get:

#summary(aov.width)
#            Df Sum Sq Mean Sq F value   Pr(>F)    
#accession    2  61.35  30.675   23.09 8.51e-07 ***
#Residuals   30  39.86   1.329 

# VG <- (30.67 - 1.32)/9 
# VE <- 1.32/9
# VP <- VG + VE 
# h2 <- (VG/VP)*100
# h2

# Would be nice to write a function to calculate this straigth from the aov summary. 

# Here I calculate the heritabilities for all the different traits. There are probably simpler ways to do this, but I am a terrible coder.

#we first calculate the results of the different analysis of variance and save them

aov.width <- aov(width ~ accession, data = global_tidy)
summary(aov.width)
aov.massY <- aov(massY ~ accession, data = global_tidy)
summary(aov.massY)
aov.area <- aov(convexhull ~ accession, data = global_tidy)
summary(aov.area)
aov.depth <- aov(depth ~ accession, data = global_tidy)
summary(aov.depth)
aov.ar <- aov(ar ~ accession, data = global_tidy)
summary(aov.ar)

# we then save the summaries in a data frame.

width <- do.call(rbind.data.frame, summary(aov.width))
massY <- do.call(rbind.data.frame, summary(aov.massY))
area <- do.call(rbind.data.frame, summary(aov.area))
depth <- do.call(rbind.data.frame,summary(aov.depth))
ar <- do.call(rbind.data.frame,summary(aov.ar))

# just adding a new column with the trait name

width$trait <- c("width", "width")
massY$trait <- c("massY", "massY")
area$trait <- c("area", "area")
depth$trait <- c("depth", "depth")
ar$trait <- c("ar", "ar")

# and with this we calculate the heritabilities according to the formula above.

width <- width %>%
        mutate(h2 = (((width[1,3] - width[2,3])/9)/ 
        (((width[1,3] - width[2,3])/9) + ((width[2,3])/9)))*100)

massY <- massY %>%
        mutate(h2 = (((massY[1,3] - massY[2,3])/9)/ 
        (((massY[1,3] - massY[2,3])/9) + ((massY[2,3])/9)))*100)

area <- area %>%
        mutate(h2 = (((area[1,3] - area[2,3])/9)/ 
        (((area[1,3] - area[2,3])/9) + ((area[2,3])/9)))*100)
        
depth <- depth %>%
        mutate(h2 = (((depth[1,3] - depth[2,3])/9)/ 
        (((depth[1,3] - depth[2,3])/9) + ((depth[2,3])/9)))*100)

ar <- ar %>%
        mutate(h2 = (((ar[1,3] - ar[2,3])/9)/ 
        (((ar[1,3] - ar[2,3])/9) + ((ar[2,3])/9)))*100)

all_traits <- rbind(width, massY, area, depth, ar)

# we then finally create a table with the traits of all the heritabilities.

h2 <- all_traits %>%
        filter(Df == 2) %>%
        select(trait, h2)


# there has to be an easiest way to do this. 

