---
title: "figure_3"
author: "Rubén Rellán-Ávarez", "Guillaume Lobet"
output: html_document
---

```{r}

library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(magrittr)
library(gridExtra)
library(rmarkdown)
library(shapes)
library(geomorph)


```




First let´s take a look at the basic global descriptors

```{r}
global <- read.csv("data/high-root-data-global.csv")

# some data tyding 

global_tidy <- global %>%
        separate(image, into = c("experiment", "das", "accession", "plant", 
                                 "aq"), sep = "-")

ggplot(global_tidy, aes(x = accession, y = width, colour = accession)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "accession", y = "Width (cm)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()

ggsave("width.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

aov.width <- aov(width ~ accession, data = global_tidy)
summary(aov.width)

# area

ggplot(global_tidy, aes(x = accession, y = convexhull, colour = accession)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "accession", y = "Root Area (cm2)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("area.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

aov.area <- aov(convexhull ~ accession, data = global_tidy)
summary(aov.area)

bay_sha <- global_tidy %>%
        filter (accession == "bay" | accession == "sha"),
wilcox.test(area ~ accession, data = bay_sha)

bay_col <- global_tidy %>%
        filter (accession == "bay" | accession == "col") 
wilcox.test(area ~ accession, data = bay_col)

sha_col <- global_tidy %>%
        filter (accession == "sha" | accession == "col") 
wilcox.test(area ~ accession, data = sha_col)

# shoot area

ggplot(global_tidy, aes(x = accession, y = shoot_area, colour = accession)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "accession", y = "Shoot Area (cm2)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("shoot_area.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

aov.shoot_area <- aov(shoot_area ~ accession, data = global_tidy)
summary(aov.shoot_area)

# depth

ggplot(global_tidy, aes(x = accession, y = depth, colour = accession)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "accession ", y = "Depth (cm)") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("depth.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

aov.depth <- aov(depth ~ accession, data = global_tidy)
summary(aov.depth)

# Aspect ratio

ggplot(global_tidy, aes(x = accession, y = ar, colour = accession)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "accession ", y = "Depth/width") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("ar.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

aov.ar <- aov(ar ~ accession, data = global_tidy)
summary(aov.ar)

# wilcox test

bay_sha <- global_tidy %>%
        filter (accession == "bay" | accession == "sha"),
wilcox.test(ar ~ accession, data = bay_sha)

bay_col <- global_tidy %>%
        filter (accession == "bay" | accession == "col") 
wilcox.test(ar ~ accession, data = bay_col)

sha_col <- global_tidy %>%
        filter (accession == "sha" | accession == "col") 
wilcox.test(ar ~ accession, data = sha_col)

 

# MasssY

ggplot(global_tidy, aes(x = accession, y = massY, colour = accession)) +
        geom_boxplot() +
        geom_point(size = 3, position=position_jitter()) +
        labs(x = "accession ", y = "massY") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw
        ggsave("massY.pdf", height = 8, width = 8, units = "cm", 
         useDingbats = FALSE)

aov.massY <- aov(massY ~ accession, data = global_tidy)
summary(aov.massY)

# wilcox test

bay_sha <- global_tidy %>%
        filter (accession == "bay" | accession == "sha"),
wilcox.test(massY ~ accession, data = bay_sha)

bay_col <- global_tidy %>%
        filter (accession == "bay" | accession == "col") 
wilcox.test(massY ~ accession, data = bay_col)

sha_col <- global_tidy %>%
        filter (accession == "sha" | accession == "col") 
wilcox.test(massY ~ accession, data = sha_col)

```

I will now take a look at root-shoots correlations.

```{r}

# shoot/root correlations (areas)

ggplot(global_tidy, aes(x = convexhull, y = shoot_area, colour = accession)) +
        geom_point(size = 3) +
        labs(x = "Root Area", y = "Shoot Area") +
        facet_wrap(~ accession, scales = "free") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("root-shoot-correlation.pdf", height = 16, width = 16, units = 
                       "cm",useDingbats = FALSE)

```


Now lets take a look at the directionality 


```{r}
dir <- read.csv("data/high-root-data-dir.csv")
dir$angle <- abs(dir$angle)

dir_tidy <- dir %>%
        separate(image, into = c("experiment", "das", "accession", "plant", "aq"), sep = "-")

ggplot(dir_tidy, aes(x = angle, y = count, colour = accession)) +
       facet_grid(~ accession) +
        geom_smooth()+
        geom_point(size = 3, alpha = 0.5) +
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        theme_bw()
        ggsave("directionality.pdf", height = 8, width = 12, units = 
                       "cm",useDingbats = FALSE) 

ggplot(dir_tidy, aes(x = angle, y = count, colour = accession)) +
        geom_point(alpha = 0.5) +
        geom_smooth()+
        facet_grid(~ accession) +
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

# separated by experiments 
dir_tidy %>%
        filter(experiment == 50) %>%
ggplot(aes(x = angle, y = count, shape = experiment)) +
       facet_grid(plant ~ accession, scales = "free") +
        geom_point(size = 2, alpha = 0.2) +
        ylim(0, 2.5) +
        labs(x = "Angle", y = "count") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality-50.pdf", height = 16, width = 12, units = 
                       "cm",useDingbats = FALSE)

dir_tidy %>%
        filter(experiment == 59) %>%
ggplot(aes(x = angle, y = count, shape = experiment)) +
       facet_grid(plant ~ accession, scales = "free") +
        geom_point(size = 2, alpha = 0.2) +
        labs(x = "Angle", y = "count") +
         ylim(0, 2.5) +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
        ggsave("directionality-59.pdf", height = 16, width = 12, units = 
                       "cm",useDingbats = FALSE)

```

Now, lets work on the heritability:

```{r}

# Example for aspect to ratio:

aov.ar <- aov(ar ~ accession, data = global_tidy)
summary(aov.ar)

# According to Olivier Loudet, the genetic variance is estimated by:

VG = (Var'Genotype' - Var'Residuals') / n

# where n is the (mean) number of replicate measurements for each genotype in the experiment. Not sure what this could be in our case.

# The 'Residuals' are a good estimator of the environmental variance in the experiment: 

VE = Var"Residuals" / n

# Total phenotipic variance would then be:

VP (Total Phenotypic variance) = VG + VE

# and therefore the (broad sense) heritability would be:

h2 = VG / VP

# So in our case for AR, we get:

# summary(aov.ar)
#            Df Sum Sq Mean Sq F value   Pr(>F)    
# accession    2  8.581   4.291   45.16 8.96e-10 ***
# Residuals   30  2.850   0.095 

# VG <- (30.67 - 1.32)/9 
# VE <- 1.32/9
# VP <- VG + VE 
# h2 <- (VG/VP)*100
# h2

# Would be nice to write a function to calculate this straigth from the aov summary. 

# Lets save the values of the summary for the differen aov

width <- do.call(rbind.data.frame, summary(aov.width))
massY <- do.call(rbind.data.frame, summary(aov.massY))
area <- do.call(rbind.data.frame, summary(aov.area))
depth <- do.call(rbind.data.frame,summary(aov.depth))
ar <- do.call(rbind.data.frame,summary(aov.ar))

width$trait <- c("width", "width")
massY$trait <- c("massY", "massY")
area$trait <- c("area", "area")
depth$trait <- c("depth", "depth")
ar$trait <- c("ar", "ar")


width <- width %>%
        mutate(h2 = (((width[1,3] - width[2,3])/9)/ 
        (((width[1,3] - width[2,3])/9) + ((width[2,3])/9)))*100)

massY <- massY %>%
        mutate(h2 = (((massY[1,3] - massY[2,3])/9)/ 
        (((massY[1,3] - massY[2,3])/9) + ((massY[2,3])/9)))*100)

area <- area %>%
        mutate(h2 = (((area[1,3] - area[2,3])/9)/ 
        (((area[1,3] - area[2,3])/9) + ((area[2,3])/9)))*100)
        
depth <- depth %>%
        mutate(h2 = (((depth[1,3] - depth[2,3])/9)/ 
        (((depth[1,3] - depth[2,3])/9) + ((depth[2,3])/9)))*100)

ar <- ar %>%
        mutate(h2 = (((ar[1,3] - ar[2,3])/9)/ 
        (((ar[1,3] - ar[2,3])/9) + ((ar[2,3])/9)))*100)

all_traits <- rbind(width, massY, area, depth, ar)

h2 <- all_traits %>%
        filter(Df == 2) %>%
        select(trait, h2)


# there has to be an easiest way to do this. 



dir2 <- read.csv("data/high-root-data-dir.csv")

dir_tidy2 <- dir2 %>%
        separate(image, into = c("experiment", "das", "accession", "plant", 
                                 "aq"), sep = "-")
dir_tidy3 <- dir_tidy2 %>%
        group_by(accession, plant, angle) %>%
        mutate(count = mean(count))
                       

write.csv(dir_tidy3, file="dir_tidy")


PCA_data <- read.csv("data/dir_tidy_wide_PCA_variable_info.csv")
PCA_data <- arrange(PCA_data, loadings.PC1)

PCA_data %>%
        ggplot(aes(x = angle, y = loadings.PC1 )) +
        geom_point(size = 3, alpha = 0.6) +
        labs(x = "Angle", y = "loadings") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
              ggsave("PCA_loadings.pdf", height = 10, width = 10, units = 
                       "cm",useDingbats = FALSE)


ANOVA_data <- read.csv("data/dir_tidy_wide_ANOVA.csv")

ANOVA_data %>%
        ggplot(aes(x = angles, y = p.values.accession_adjusted.p.values )) +
        geom_point(size = 3, alpha = 0.6) +
        labs(x = "Angle", y = "P-values") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
              ggsave("P-values-.pdf", height = 10, width = 10, units = 
                       "cm",useDingbats = FALSE)


ANOVA_data %>%
        ggplot(aes(x = angles, y = p.values.accession_q.values)) +
        geom_point(size = 3, alpha = 0.6) +
        labs(x = "Angle", y = "P-values") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
              ggsave("PCA_loadings.pdf", height = 10, width = 10, units = 
                       "cm",useDingbats = FALSE)

ANOVA_data %>%
        ggplot(aes(x = angles, y = p.values.accession_q.values)) +
        geom_point(size = 3, alpha = 0.6) +
        labs(x = "Angle", y = "P-values") +
              theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
              ggsave("PCA_loadings.pdf", height = 10, width = 10, units = 
                       "cm",useDingbats = FALSE)
      
```

Directionality using the new local analysis.

```{r}

dir_local <- read.csv("data/root-data-local.csv")
dir_local$angle <- abs(dir_local$angle)

dir_local_tidy <- dir_local %>%
        separate(image, into = c("experiment", "das", "accession", "plant", 
                                 "aq", sep="-"))

dir_local_tidy[,6] <- NULL

ggplot(dir_local_tidy, aes(x=length, fill = accession))+
  geom_histogram(alpha=.5, binwidth = 5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())



ggplot(dir_local_tidy, aes(x=angle, fill = accession))+
  geom_density(alpha=.5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

ggplot(dir_local_tidy, aes(x=x, y= desc(y), colour = angle ))+
  geom_point(alpha=0.5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

```

I was getting very small segments with the high contrast option so I run it
on low contrast:

```{r}

dir_local_low <- read.csv("data/root-data-low-local.csv")
dir_local_low$angle <- abs(dir_local_low$angle)

dir_local_low_tidy <- dir_local_low %>%
        separate(image, into = c("experiment", "das", "accession", "plant", 
                                 "aq", sep="-"))

dir_local_low_tidy[,6] <- NULL

ggplot(dir_local_low_tidy, aes(x=length, fill = accession))+
  geom_histogram(alpha=.5, binwidth = 5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())



ggplot(dir_local_low_tidy, aes(x=angle, fill = accession))+
  geom_density(alpha=.5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

ggplot(dir_local_low_tidy, aes(x=x, y= desc(y), colour = angle ))+
  geom_point(alpha=0.5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

```

testing the Kolmogorov-Smirnoff test:

```{r}

dir_local_low <- read.csv("data/root-data-low-local.csv")
dir_local_low$angle <- abs(dir_local_low$angle)

dir_local_low_tidy <- dir_local_low %>%
        separate(image, into = c("experiment", "das", "accession", "plant", 
                                 "aq", sep="-"))

dir_local_low_tidy[,6] <- NULL

ggplot(dir_local_low_tidy, aes(x=length, fill = accession))+
  geom_histogram(alpha=.5, binwidth = 5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())



ggplot(dir_local_low_tidy, aes(x=angle, fill = accession))+
  geom_density(alpha=.5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

ggplot(dir_local_low_tidy, aes(x=x, y= desc(y), colour = angle ))+
  geom_point(alpha=0.5) +
  facet_grid (~ accession) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank())

```

Shape analysis

```{r}


#read your tps file with landmarkcoordinates, create 2D array
myData <- readland.tps("data/root-data-shape.tps", specID = "ID")

#read your csv with classifiers
temp <- read.csv("data/root-data-shape.csv")
classifier <- data.frame(name=temp$image)
for(i in 1:nrow(classifier)){
        temp <- strsplit(as.character(classifier$name[i]), "-")    
        classifier$genotype[i] <- temp[[1]][3]
        classifier$treatment[i] <- temp[[1]][4]
}

# Make the procust analysis
proc <- procGPA(myData)

# Get the values from the PCA
classifier$PC1 <- proc$scores[,1]
classifier$PC2 <- proc$scores[,2]
classifier$PC3 <- proc$scores[,3]

# get the mean shape
shape <- data.frame(x = proc$mshape[,1], y = -proc$mshape[,2])

# Get the deviations
rotPC1 <-  proc$pcar[, 1]
rotPC2 <-  proc$pcar[, 2]
rotPC3 <-  proc$pcar[, 3]
sdPC1 <- proc$pcasd[1]
sdPC2 <- proc$pcasd[2]
sdPC3 <- proc$pcasd[3]

# Base plot
plot <- ggplot(data = shape, aes(x=x, y=y)) 
for(i in 1:proc$n){
        temp <- data.frame(x=proc$rotated[,1,i], y=-proc$rotated[,2,i])
        plot <- plot + geom_polygon(data=temp, aes(x=x, y=y), fill="transparent",colour="#00000030") 
}
plot <- plot + geom_polygon(fill="transparent", size=2, colour="#00000080")  + theme_bw()
devs <- c(-3, 3)
col <- c("#FF605D", "#00B94E")

### PC1
plot1 <- plot + ggtitle(paste("PC1 [",round(proc$percent[1],2),"%]"))
for(i in 1:length(devs)){
        fig <- proc$mshape + devs[i] * rotPC1 * sdPC1
        k <- length(proc$mshape[,1])
        temp <- data.frame(x= fig[1:k], y=-fig[(k + 1):(2 * k)])
        plot1 <- plot1 + geom_polygon(data=temp, aes(x=x, y=y), fill="transparent",colour=col[i], size=1) 
}

### PC2
plot2 <- plot + ggtitle(paste("PC2 [",round(proc$percent[2],2),"%]"))
for(i in 1:length(devs)){
        fig <- proc$mshape + devs[i] * rotPC2 * sdPC2
        k <- length(proc$mshape)/2
        temp <- data.frame(x= fig[1:k], y=-fig[(k + 1):(2 * k)])
        plot2 <- plot2 + geom_polygon(data=temp, aes(x=x, y=y), fill="transparent",colour=col[i], size=1) 
}

### PC3
plot3 <- plot + ggtitle(paste("PC3 [",round(proc$percent[3],2),"%]"))
for(i in 1:length(devs)){
        fig <- proc$mshape + devs[i] * rotPC3 * sdPC3
        k <- length(proc$mshape)/2
        temp <- data.frame(x= fig[1:k], y=-fig[(k + 1):(2 * k)])
        plot3 <- plot3 + geom_polygon(data=temp, aes(x=x, y=y), fill="transparent",colour=col[i], size=1) 
}

pcaplot <- qplot(data = classifier, x = PC1, y = PC2, colour = genotype, main="Shape Analysis") + 
        stat_ellipse(level = 0.95, size=1, fill = genotype) + 
        coord_fixed(ratio = 1) +
        theme_bw() + 
        scale_x_continuous(name=paste("PC1 (",round(proc$percent[1],2),"%)",sep="")) +
        scale_y_continuous(name=paste("PC2 (",round(proc$percent[2],2),"%)",sep="")) +
        ggsave("shape_PCA.pdf", height = 10, width = 12, units = "cm", 
         useDingbats = FALSE)

grid.arrange(arrangeGrob(plot1, plot2, ncol=1), pcaplot, ncol=2)
grid.arrange(plot1, plot2, pcaplot, ncol=3) 
g <- arrangeGrob(plot1, plot2, pcaplot,  ncol=3, widths=c(1, 1, 1.75))
  ggsave("shape.pdf", g,  height = 8, width = 18, units = "cm", 
         useDingbats = FALSE)

```

ggplot(global_tidy, aes(x = circularity, y = ar, colour = accession)) +
        geom_point(size = 3) +
        theme_bw()
        
ggplot(global_tidy, aes(x = round, y = ar, colour = accession)) +
        geom_point(size = 3) +
        theme_bw()
